DOpElib: A Goal Oriented Software Library for Computing PDEs and Optimization Problems

### 2012-10-22 Paper_May_2012/main.tex
update of part 2.1.2

In order to solve this problem, i.e., calculate $J(u)$, we need to discretize 
the equation, in \dope, we are\marginpar{here using?} using finite elements to this end. This means we need to pick a 
subdivision $\mathcal T_h$, for instance into quadrilaterals,
of the domain $\Omega$ and a finite element. In the concrete case we could, 
for instance, take the $\mathcal Q_2/\mathcal Q_1$ Taylor-Hood element.
As it is clear, that the initialization of the unknowns, the local shape 
functions, removal of hanging node constraints, etc. always follows the 
same procedure, we define an object which takes the finite element (FE)
and the triangulation and takes care of all other things related with 
handling the finite element space $V_h$. We will refer, to 
this as the \texttt{SpaceTimeHandler}. It is then clear, that the simplest 
form of the constructor should have the following interface:
\begin{lstlisting}
  template<typename Triangulation, typename FE>
      void SpaceTimeHandler(Triangulation &, FE);
\end{lstlisting}
Next, one needs to solve the discrete equation
\begin{align}\label{eq:discrete_equation}
a(u_h)(\phi_h) = f(\phi_h) \quad \forall \phi \in V_h.
\end{align}
To this end one can employ a newton type method with some globalization 
like line-search. To do so, one needs to calculate the residual of the 
equation, as well as solving linearized equations for the calculation of 
the newton method and then employing some kind of globalization strategy.
This again will usually require the evaluation of the residual.

The evaluation of the residual will require the calculation of the integrals 
involved in the weak formulation of the problem. For example 
the $i$-th component of the residual, with respect to the given FE-basis 
$\phi_{h,i}$, will read as follows
\[
 a(u_h)(\phi_{h,i}) - f(\phi_{h,i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u, \nabla \phi_i\,dx + \ldots
\]
Again, we see that, except for the local element integrals, all 
calculations are independent from the concrete form of the PDE. 
Hence we define an object \texttt{Integrator}
that offers a method to calculate the residual given a user-defined 
object \texttt{pde} that implements the local element integrals, i.e.,
\begin{lstlisting}
  template<typename PROBLEM,typename VECTOR>
      void Integrator
             ::ComputeNonlinearResidual(PROBLEM& pde, 
                                        VECTOR & residual);
\end{lstlisting}
where \texttt{VECTOR} is one of the possible vector types given by \deal{}.
It is clear that we need to ensure that the integrator object has access
to the information stored in the \texttt{SpaceTimeHandler}
to correctly loop over all elements.
To solve the linearized systems, again, all that is left is the calculation
of the matrix corresponding to the linearized equation. To calculate this,
again we can define a method in the integrator that does exactly this provided
we have a user defined object \texttt{pde} that can evaluate the local 
integrals needed for assembling the matrix, i.e.,
\begin{lstlisting}
  template<typename PROBLEM,typename MATRIX>
      void Integrator
             ::ComputeMatrix(PROBLEM& pde, 
                             MATRIX & system_matrix);
\end{lstlisting}
where now \texttt{MATRIX} is one of the matrix types supported by \deal{}.
With this we can now define a nonlinear solver \texttt{NewtonSolver} 
for the solution of the discrete problem requiring the existence of the 
following function
\begin{lstlisting}
  template<typename PROBLEM, typename VECTOR>
      void NewtonSolver
             ::NonlinearSolve(PROBLEM& pde, 
                              VECTOR & solution);
\end{lstlisting}

@@

We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and the functional $J$ is defined as [TODO: Definition von J. Nutze Benchmark case oder lenkt das zu sehr ab?]. Additionally, for given inflow $v_{\operatorname{in}}$ we set  boundary conditions $
v = v_{\operatorname{in}} \text{ on } \Gamma_{in}\text{ and }v = 0  \text{ on } \Gamma_{wall}$.
\marginpar{Das Beispiel hier besser nochmal zusammen schreiben?}

\begin{figure}[h]
\centering
\resizebox{0.5\textwidth}{!}{\input{Pictures/example_ns}}
\caption{Flow around a cylinder with 
circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
\label{fig:example_ns}
\end{figure}

We split our program into the files \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}. The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional. We give a brief explanation on every file in the following.

@@

After the usual preamble, in which we include the necessary header-files from \deal{}, \dope{} and our own problem (i.e. \texttt{localfunctional.h} and \texttt{localpde.h}), we start by defining an instance of a \texttt{ParameterReader}-object, which handles run-time parameter.
\begin{lstlisting}
   ParameterReader pr;
\end{lstlisting}
In a text-file we might specify some problem-related parameters (for instance the viscosity $\nu$), but also general parameters for (non)-linear solvers and the output of computed data. 
After that, we create a triangulation named \texttt{triangulation}. This is done exactly as in \deal{}, so we can choose between several possibilities: We could explicitly create a grid by providing a list of vertices and cells, read in a triangulation via \texttt{dealii::GridIn}-class or use the \texttt{dealii::GridGenerator}, which offers common domains like hypercubes or hyper-balls.

[code, Grid einlesen im Benchmark-Fall, ansonsten GridGenerator something. Hier auch auf Randfarben kurz eingehen.]

Next, we create  quadrature rules (for the evaluation of integrals over cells and faces) and a finite element appropriate for the PDE at hand. We choose a gaussian quadrature of order three and the previosly mentioned Tayler-Hood element. \marginpar{Wie klar machen, dass vieles von deal.II uebernommen und nicht selbst geschrieben?}
\begin{lstlisting}
  QGauss<2> quad_formula(3);
  QGauss<1> face_quad_formula(3);
  FESystem<2> finite_element(dealii::FE_Q<2>(2), 2,
                             dealii::FE_Q(1), 1);
\end{lstlisting}
%IntegratorDataContainer<VECTOR, 2> idc(quad_formula,
%                                         face_quad_formula);
%The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
Then, the \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play.
\begin{lstlisting}
LocalPDE<VECTOR> local_pde;
LocalFunctional<VECTOR> local_functional;
\end{lstlisting}
These hold the information of the PDE and functional under consideration. Further explication is given below because
the definition of this classes is done spererately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.

In the next step, we define an object of type \texttt{SpaceTimeHandler}, which handles all the things connected with the triangulation as well as the finite element space. This class is also capable of organizing degrees of freedom in time-dependent problems, hence the name. 
\begin{lstlisting}
  SpaceTimeHandler<FE> dof_handler(triangulation, finite_element);
\end{lstlisting}
There exist different versions of this class for forward as well as optimization problems.

Now we have everything at hand to create an instantiation of type \texttt{PDEProblemContainer}. This basically represents the disrectized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization. First, we give the constructor the \texttt{local\_pde} and the \texttt{dof\_handler}.
\begin{lstlisting}
  PDEProblemContainer<LocalPDE> prob_container(local_pde,
                                                  dof_handler);
\end{lstlisting}
After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor').
\begin{lstlisting}
  prob_container.AddFunctional(&local_functional);
  prob_container.SetBoundaryFunctionalColors(1);
\end{lstlisting}
Next, we incorporate the dirichlet data. First, we define two functions representing the dirichlet boundary conditions. These are given to an interface\marginpar{SimpleDirichletData weglassen?} class \texttt{SimpleDirichletData}. 
\begin{lstlisting}
  ZeroFunction<2> zf;
  InflowFunction<2> if;
  SimpleDirichletData<VECTOR, 2> dd_zf(zf);
  SimpleDirichletData<VECTOR, 2> dd_if(if);
\end{lstlisting}
Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the dirichlet conditions.
\begin{lstlisting}
  std::vector<bool> comp_mask(3);

  comp_mask[0] = true;
  comp_mask[1] = true;
  comp_mask[2] = false;

  prob_container.SetDirichletBoundaryColors(0, comp_mask, &dd_if);
  prob_container.SetDirichletBoundaryColors(2, comp_mask, &dd_zf);
  prob_container.SetDirichletBoundaryColors(3, comp_mask, &dd_zf);
\end{lstlisting}
Now we create the object which steers the whole solution process, the previosly mentioned \texttt{StatPDEProblem}. It takes care of solving the discretized equation and evaluating the given functionals. The class depends actually on some template-parameters, so for the sake of abbreviation, we defined a typedef at the beginning of the \texttt{main.cc}-file.
\begin{lstlisting}
typedef StatPDEProblem<NEWTONSOLVER,
           PDEProblemContainer<LocalPDE>, VECTOR, 2> SSOLVER;
\end{lstlisting}
Here, \texttt{NEWTONSOLVER} itself is an abbreviation of
\begin{lstlisting}
typedef NewtonSolver<LINEARSOLVER, VECTOR, 2> NEWTONSOLVER;
\end{lstlisting}
The first template parameter of \texttt{NewtonSolver} describes which linear solver we use in this example. We opted for a direct solver, so: 
\begin{lstlisting}
typedef DirectLinearSolverWithMatrix<MATRIX,
            VECTOR, 2> LINEARSOLVER;
\end{lstlisting}
\marginpar{Kein IDC, dafuer quadformeln eingefuegt.}
With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
\begin{lstlisting}
  SSOLVER solver(&prob_container, pr, quad_formula, face_quad_formula);
\end{lstlisting}
\marginpar{outputhandler erlaeutern?}
%Finally, the output handler is initialized via
%\begin{lstlisting}
%  P.RegisterOutputHandler(&out);
%  solver.RegisterOutputHandler(&out);
%\end{lstlisting}
Finally, the programm is started with:
\begin{lstlisting}
  solver.ReInit();
  solver.ComputeReducedFunctionals();
\end{lstlisting}

@@

\marginpar{Rekapitulierung des Newton-Verfahrens?}

Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h,i}|1\leq i \leq N}$ be a given FE-Basis. As stated in [sectionref], to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e. the vector
\begin{align}\label{eq:discrete_residual}
\left(a(u_h)(\phi_{h,i})-(f,\phi_{h,i})\right)_{i=1}^N,
\end{align}
and the Jacobi-matrix of the weak form, i.e.
\begin{align}\label{eq:discrete_matrix}
\left(\nabla a(u_h)(\phi_{h,j},\phi_{h,i})\right)_{i,j=1}^N.
\end{align}
Now all we have to supply here are the cell-wise contributions to these terms, the \texttt{Integrator} takes care of assembling \eqref{eq:discrete_residual} and \eqref{eq:discrete_matrix}.

@@

\begin{lstlisting}
  void CellEquation(const CellDataContainer<2>& cdc,
                    dealii::Vector<double> &local_cell_vector)
    {
      const auto& fe_values = cdc.GetFEValuesState();
      unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
      unsigned int n_q_points = cdc.GetNQPoints();
\end{lstlisting}
 The \texttt{CellDataContainer} contains all the information we need to compute the cell-contributions (things like number of quadrature points, number of degrees of freedom on the cell, the \texttt{dealii::FEValues}-objects, etc.).
%  \begin{lstlisting}
%      _uvalues.resize(n_q_points,Vector<double>(3));
%      _ugrads.resize(n_q_points,vector<Tensor<1,2> >(3));
%      cdc.GetValuesState("last_newton_solution", _uvalues);
%      cdc.GetGradsState("last_newton_solution", _ugrads);
%  \end{lstlisting}
  \begin{lstlisting}
     cdc.GetValuesState("last_newton_solution", _uvalues);
     cdc.GetGradsState("last_newton_solution", _ugrads);
 \end{lstlisting}
\texttt{\_uvalues} and \texttt{\_ugrads} are of type \texttt{std::vector<...>}, with $\ldots$ symbolizing the appropriate data type. In these two lines, we load the (gradient of the) solution of the last newton iteration in all the quadrature points on the cell into the two \texttt{std::vector} objects.

In the next linex, we loop over the quadrature points and the number of dofs of the cell. In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ on the cell with the help of the previosly selected quadrature formula. 
%     const FEValuesExtractors::Vector velocities(0);
%     const FEValuesExtractors::Scalar pressure(2); 
 \begin{lstlisting}
    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
    {
      Tensor<2,2> vgrads;
      vgrads[0][0] = _ugrads[q_point][0][0]; 
      //etc.
      const Tensor<2,2> phi_i_grads_v = 
          state_fe_values[velocities].gradient (i, q_point);
      for (unsigned int i=0;i<n_dofs_per_cell;i++)
        local_cell_vector(i) += ( _nu 
                                *_vgrads[q_point] * phi_i_grads_v 
                                + //all the other terms)
                                * state_fe_values.JxW(q_point);
\end{lstlisting}

% [CellEquation]
% 
% \begin{lstlisting}
%   void CellRightHandSide(const CellDataContainer<2>& cdc,
%                          dealii::Vector<double> &local_cell_vector,
%                          double scale);
% \end{lstlisting}
% 
% \begin{lstlisting}
%   void CellMatrix(const CellDataContainer<2>& cdc,
%                   dealii::FullMatrix<double> &local_entry_matrix);
% \end{lstlisting}
% The PDE under consideration is nonlinear, so let us briefly recapitulate manner such that for solving 
% the Laplace problem the following nonlinear step is necessary to recapitulate:
% Find $\delta u\in V$:
% \[
% (\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
% \]
% and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
% Therefore, the implementation reads for the right hand side:
% \begin{lstlisting}
%    void CellEquation(...)
%     {
%       // implement (\nabla u, \nabla \phi) 
%     }
% \end{lstlisting}
% Specifically, the inner loops reads
% \begin{lstlisting}
%    for (all quadrature points per cell q)
%       {
%         Tensor<2, 2> ugrads; // solution vector
%  
% 	for (all dofs per cell i)
%         {
% 	  const Tensor<2, 2> phi_i_grads_u; // test function
% 
%           local_cell_vector(i) += scale * 
%             scalar_product(ugrads, phi_i_grads_u) *
% 	    state_fe_values.JxW(q_point);
%         }
%       }
% \end{lstlisting}
% For the left hand side, we need to implement the Jacobian matrix:
% \begin{lstlisting}
%    void CellMatrix(...)
%     {
%       // implement  (\nabla \delta u, \nabla \phi)
%     }
% \end{lstlisting}
% Specifically,
% \begin{lstlisting}
%       for (all quadrature points per cell q)      
%         for (all dofs per cell i)    
%           for (all dofs per cell j)
%           {
%             local_entry_matrix(i, j) += 
%               scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
%               state_fe_values.JxW(q_point);
%           }
% \end{lstlisting}
% The integration of the equation for each DoF and each quadrature point on each
% cell is then done 
% in the \texttt{templates/integrator.h} function. Finally, as usually in 
% finite element computations, this leads to a discrete system. Since
% we treat each problem with a Newton method, within each Newton step, the 
% following linear system must be solved:
% \[
% A\delta u = b
% \]
% in which $A$ is specified by \texttt{CellMatrix(...)} and 
% $b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
% present Newton step:
% \[
% u^n = u^{n-1} + \omega \delta u
% \]
% with a line search parameter $\omega\in (0,1]$.

\paragraph{\texttt{localfunctional.h}}
[TODO Christian]

>>>>

In order to solve this problem, i.e., calculate $J(u)$, we need to discretize 
the equation. In \dope, we are using finite elements to do this. This means we need to pick a 
subdivision $\mathcal T_h$, for instance into quadrilaterals,
of the domain $\Omega$ and a finite element. In the concrete case we could, 
for instance, take the $\mathcal Q_2/\mathcal Q_1$ Taylor-Hood element.
As it is clear, that the initialization of the unknowns, the local shape 
functions, removal of hanging node constraints, etc. always follows the 
same procedure, we define an object which takes the finite element (FE)
and the triangulation and takes care of all other things related with 
handling the finite element space $V_h$. We will refer, to 
this as the \texttt{SpaceTimeHandler}. It is then clear, that the simplest 
form of the constructor should have the following interface:
\begin{lstlisting}
  template<typename Triangulation, typename FE>
      void SpaceTimeHandler(Triangulation &, FE);
\end{lstlisting}
Next, one needs to solve the discrete equation
\begin{align}\label{eq:discrete_equation}
a(u_h)(\phi_h) = f(\phi_h) \quad \forall \phi \in V_h.
\end{align}
To this end one can employ a newton type method with some globalization 
like line-search. To do so, one needs to calculate the residual of the 
equation, as well as solving linearized equations for the calculation of 
the newton method and then employing some kind of globalization strategy.
This again will usually require the evaluation of the residual.

The evaluation of the residual will require the calculation of the integrals 
involved in the weak formulation of the problem. For example 
the $i$-th component of the residual, with respect to the given FE-basis 
$\phi_{h,i}$, will read as follows
\begin{align}\label{eq:residual_vector}
 a(u_h)(\phi_{h,i}) - f(\phi_{h,i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u, \nabla \phi_i\,dx + \ldots
 \end{align}
Again, we see that, except for the local element integrals, all 
calculations are independent from the concrete form of the PDE. 
Hence we define an object \texttt{Integrator}
that offers a method to calculate the residual given a user-defined 
object \texttt{pde} that implements the local element integrals, i.e.,
\begin{lstlisting}
  template<typename PROBLEM,typename VECTOR>
      void Integrator
             ::ComputeNonlinearResidual(PROBLEM& pde, 
                                        VECTOR & residual);
\end{lstlisting}
where \texttt{VECTOR} is one of the possible vector types given by \deal{}.
It is clear that we need to ensure that the integrator object has access
to the information stored in the \texttt{SpaceTimeHandler}
to correctly loop over all elements.
To solve the linearized systems, again, all that is left is the calculation
of the matrix corresponding to the linearized equation. To calculate this,
again we can define a method in the integrator that does exactly this provided
we have a user defined object \texttt{pde} that can evaluate the local 
integrals needed for assembling the matrix, i.e.,
\begin{lstlisting}
  template<typename PROBLEM,typename MATRIX>
      void Integrator
             ::ComputeMatrix(PROBLEM& pde, 
                             MATRIX & system_matrix);
\end{lstlisting}
where now \texttt{MATRIX} is one of the matrix types supported by \deal{}.
With this we can now define a nonlinear solver \texttt{NewtonSolver} 
for the solution of the discrete problem requiring the existence of the 
following function
\begin{lstlisting}
  template<typename PROBLEM, typename VECTOR>
      void NewtonSolver
             ::NonlinearSolve(PROBLEM& pde, 
                              VECTOR & solution);
\end{lstlisting}

@@

We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle. We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$. This is the Benchmark-Case 2D-1 from \cite{TuSchae96}, see there for further details.
%Additionally, for given inflow $v_{\operatorname{in}}$ we set  boundary conditions 
%v = v_{\operatorname{in}} \text{ on } \Gamma_{in}\text{ and }v = 0  \text{ on } \Gamma_{wall}.
\begin{figure}[h]
\centering
\resizebox{0.5\textwidth}{!}{\input{Pictures/example_ns}}
\caption{Flow around a cylinder with 
circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
\label{fig:example_ns}
\end{figure}

We split the program into the parts \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}. The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional. We give a brief explanation on every file in the following.

@@

After the usual preamble, in which we include the necessary header-files from \deal{}, \dope{} and our own problem (i.e. \texttt{localfunctional.h} and \texttt{localpde.h}), we start by defining an instance of a \texttt{ParameterReader}-object, which handles run-time parameter.
\begin{lstlisting}
   ParameterReader pr;
\end{lstlisting}
In a text-file we might specify some problem-related parameters (for instance the viscosity $\nu$), but also general parameters for (non)-linear solvers and the output of computed data. 
After that, we create a triangulation of the domain $\Omega$ named \texttt{triangulation}.
%This is done exactly as in \deal{}, so we can choose between several possibilities: We could explicitly create a grid by providing a list of vertices and cells, read in a triangulation via \texttt{dealii::GridIn}-class or use the \texttt{dealii::GridGenerator}, which offers common domains like hypercubes or hyper-balls.

Next, we create some quadrature rules (for the evaluation of integrals over cells and faces) and a finite element appropriate for the PDE at hand. We choose a gaussian quadrature of order three and the previosly mentioned Tayler-Hood element. \marginpar{Wie klar machen, dass vieles von deal.II uebernommen und nicht selbst geschrieben?}
\begin{lstlisting}
  QGauss<2> quad_formula(3);
  QGauss<1> face_quad_formula(3);
  FESystem<2> finite_element(dealii::FE_Q<2>(2), 2,
                             dealii::FE_Q(1), 1);
\end{lstlisting}
%IntegratorDataContainer<VECTOR, 2> idc(quad_formula,
%                                         face_quad_formula);
%The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
Then, the \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play.
\begin{lstlisting}
LocalPDE<VECTOR> local_pde;
LocalFunctional<VECTOR> local_functional;
\end{lstlisting}
These hold the information of the PDE and functional under consideration. Further explication is given below because
the definition of this classes is done spererately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.

In the next step, we define an object of type \texttt{SpaceTimeHandler}, which handles all the things connected with the triangulation as well as the finite element space. This class is also capable of organizing degrees of freedom in time-dependent problems, hence the name. 
\begin{lstlisting}
  SpaceTimeHandler<TRIANGULATION, FE> dof_handler(triangulation,
                                                  finite_element);
\end{lstlisting}
There exist different versions of this class for forward as well as optimization problems.

Now we have everything at hand to create an instantiation of type \texttt{PDEProblemContainer}. This basically represents the disrectized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization. First, we give the constructor the \texttt{local\_pde} and the \texttt{dof\_handler}.
\begin{lstlisting}
  PDEProblemContainer<LocalPDE> prob_container(local_pde,
                                               dof_handler);
\end{lstlisting}
After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor').
\begin{lstlisting}
  prob_container.AddFunctional(&local_functional);
  prob_container.SetBoundaryFunctionalColors(1);
\end{lstlisting}
Next, we incorporate the dirichlet data, which is given by the class \texttt{DirichleData}. Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the dirichlet conditions. This is done via a boundary-color (0 stands for $\Gamma_D$) and a \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
\begin{lstlisting}
  DirichletData dd;
  prob_container.SetDirichletBoundaryColors(0, comp_mask, dd);
\end{lstlisting}
Now we create the object which steers the whole solution process, the previosly mentioned \texttt{StatPDEProblem}. It takes care of solving the discretized equation and evaluating the given functionals. The class depends actually on some template-parameters, so for the sake of abbreviation, we defined a typedef at the beginning of the \texttt{main.cc}-file.
\begin{lstlisting}
typedef StatPDEProblem<NEWTONSOLVER,
           PDEProblemContainer<LocalPDE>, VECTOR, 2> SSOLVER;
\end{lstlisting}
Here, \texttt{NEWTONSOLVER} itself is an abbreviation of
\begin{lstlisting}
typedef NewtonSolver<LINEARSOLVER, VECTOR, 2> NEWTONSOLVER;
\end{lstlisting}
The first template parameter of \texttt{NewtonSolver} describes which linear solver we use in this example. We opted for a direct solver, so: 
\begin{lstlisting}
typedef DirectLinearSolverWithMatrix<MATRIX,
            VECTOR, 2> LINEARSOLVER;
\end{lstlisting}
With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
\begin{lstlisting}
  SSOLVER solver(prob_container, pr, quad_formula, face_quad_formula);
\end{lstlisting}
%Finally, the output handler is initialized via
%\begin{lstlisting}
%  P.RegisterOutputHandler(&out);
%  solver.RegisterOutputHandler(&out);
%\end{lstlisting}
Finally, the programm is started by
\begin{lstlisting}
  solver.ReInit();
  solver.ComputeReducedFunctionals();
\end{lstlisting}

@@

%\marginpar{Rekapitulierung des Newton-Verfahrens?}
Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h,i}|1\leq i \leq N}$ be a given FE-Basis. As stated in [sectionref], to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e. the vector
\begin{align}\label{eq:discrete_residual}
\left(a(u_h)(\phi_{h,i})-(f,\phi_{h,i})\right)_{i=1}^N,
\end{align}
and the Jacobi-matrix of the weak form, i.e.
\begin{align}\label{eq:discrete_matrix}
\left(\nabla a(u_h)(\phi_{h,j},\phi_{h,i})\right)_{i,j=1}^N.
\end{align}
Now all we have to supply here are the cell-wise contributions to these terms, the \texttt{Integrator} takes care of assembling \eqref{eq:discrete_residual} and \eqref{eq:discrete_matrix}.

@@

\begin{lstlisting}
  void CellEquation(const CellDataContainer<2>& cdc,
                    VECTOR &local_cell_vector)
    {
      const auto& fe_values = cdc.GetFEValuesState();
      unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
      unsigned int n_q_points = cdc.GetNQPoints();
\end{lstlisting}
 The \texttt{CellDataContainer} contains all the information we need to compute the cell-contributions (things like number of quadrature points, number of degrees of freedom on the cell, the \texttt{dealii::FEValues}-objects, etc.).
%  \begin{lstlisting}
%      _uvalues.resize(n_q_points,Vector<double>(3));
%      _ugrads.resize(n_q_points,vector<Tensor<1,2> >(3));
%      cdc.GetValuesState("last_newton_solution", _uvalues);
%      cdc.GetGradsState("last_newton_solution", _ugrads);
%  \end{lstlisting}
  \begin{lstlisting}
     cdc.GetValuesState("last_newton_solution", uvalues);
     cdc.GetGradsState("last_newton_solution", ugrads);
 \end{lstlisting}
 In these two lines, we load the (gradient of the) solution of the last newton iteration in all the quadrature points on the cell into the two vectors \texttt{uvalues} and \texttt{ugrads}.

In the next lines, we loop over the quadrature points and the number of DoFs of the cell. In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previosly selected quadrature formula. 
%     const FEValuesExtractors::Vector velocities(0);
%     const FEValuesExtractors::Scalar pressure(2); 
 \begin{lstlisting}
    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
    {
      //initialize phi_i_grads_v etc.
      Tensor<2,2> vgrads;
      vgrads[0]= ugrads[q_point][0]; 
      vgrads[1]= ugrads[q_point][1];
      for (unsigned int i=0;i<n_dofs_per_cell;i++)
        local_cell_vector(i) += ( _nu 
                                * vgrads * phi_i_grads_v 
                                + //all the other terms)
                                * state_fe_values.JxW(q_point);
    }
\end{lstlisting}
\dope{} takes now care of the distribution of the local cell contributions into  the vector of the residual \eqref{eq:residual_vector}. The methods \texttt{CellRightHandSide} and \texttt{CellMatrix} follow the same structure as 
\texttt{CellEquation}, so we do not go into the details here.
% [CellEquation]
% 
% \begin{lstlisting}
%   void CellRightHandSide(const CellDataContainer<2>& cdc,
%                          dealii::Vector<double> &local_cell_vector,
%                          double scale);
% \end{lstlisting}
% 
% \begin{lstlisting}
%   void CellMatrix(const CellDataContainer<2>& cdc,
%                   dealii::FullMatrix<double> &local_entry_matrix);
% \end{lstlisting}
% The PDE under consideration is nonlinear, so let us briefly recapitulate manner such that for solving 
% the Laplace problem the following nonlinear step is necessary to recapitulate:
% Find $\delta u\in V$:
% \[
% (\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
% \]
% and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
% Therefore, the implementation reads for the right hand side:
% \begin{lstlisting}
%    void CellEquation(...)
%     {
%       // implement (\nabla u, \nabla \phi) 
%     }
% \end{lstlisting}
% Specifically, the inner loops reads
% \begin{lstlisting}
%    for (all quadrature points per cell q)
%       {
%         Tensor<2, 2> ugrads; // solution vector
%  
% 	for (all dofs per cell i)
%         {
% 	  const Tensor<2, 2> phi_i_grads_u; // test function
% 
%           local_cell_vector(i) += scale * 
%             scalar_product(ugrads, phi_i_grads_u) *
% 	    state_fe_values.JxW(q_point);
%         }
%       }
% \end{lstlisting}
% For the left hand side, we need to implement the Jacobian matrix:
% \begin{lstlisting}
%    void CellMatrix(...)
%     {
%       // implement  (\nabla \delta u, \nabla \phi)
%     }
% \end{lstlisting}
% Specifically,
% \begin{lstlisting}
%       for (all quadrature points per cell q)      
%         for (all dofs per cell i)    
%           for (all dofs per cell j)
%           {
%             local_entry_matrix(i, j) += 
%               scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
%               state_fe_values.JxW(q_point);
%           }
% \end{lstlisting}
% The integration of the equation for each DoF and each quadrature point on each
% cell is then done 
% in the \texttt{templates/integrator.h} function. Finally, as usually in 
% finite element computations, this leads to a discrete system. Since
% we treat each problem with a Newton method, within each Newton step, the 
% following linear system must be solved:
% \[
% A\delta u = b
% \]
% in which $A$ is specified by \texttt{CellMatrix(...)} and 
% $b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
% present Newton step:
% \[
% u^n = u^{n-1} + \omega \delta u
% \]
% with a line search parameter $\omega\in (0,1]$.

\paragraph{\texttt{localfunctional.h}}
The implementation of the boundary-functional $J$ follows the outline of the computation of the cell-residual in \texttt{localpde.h}. However, there are two main differences. First, we do not have to specify the method \texttt{CellEquation}, but \texttt{BoundaryValue} instead. This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face wise (for the faces that lie on $\Gamma_{circ}$). The second difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
\begin{lstlisting}
 double
 BoundaryValue(const FaceDataContainer<2>& fdc)
  {
    //same as before, get n_q_points etc.
    fdc.GetFaceValuesState("state", ufacevalues);
    fdc.GetFaceGradsState("state", ufacegrads);
\end{lstlisting}
Notice that we get here the values of the computed solution by giving the string "state".
\begin{lstlisting}
    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
      erg += -1. * ( _nu * ufacegrads[q_point][0]
                * state_fe_face_values.normal_vector(q_point)
                + //other terms)        
                * state_fe_face_values.JxW(q_point);
    return erg;
  }
\end{lstlisting}
